---
title: "Global $CO_{2}$ Emissions in 1997"
short: "What Keeling missed all these years"
journal: "AER" # AER, AEJ, PP, JEL
month: "`r format(Sys.Date(), '%m')`"
year: "`r format(Sys.Date(), '%Y')`"
vol: 0
issue: 0
keywords:
  - Replication
  - Modern Science
author:
  - name: Finnian Meagher
    firstname: Finnian
    surname: Meagher
    email: fmeagher@berkeley.edu
    affiliation: UC Berkeley, School of Information
  - name: Kumar Narayanan
    firstname: Kumar
    surname: Narayanan
    email: kumarn@ischool.berkeley.edu
    affiliation: UC Berkeley, School of Information
  - name: Sandeep Kataria
    firstname: Sandeep 
    surname: Kataria
    email: kataria@berkeley.edu
    affiliation: UC Berkeley, School of Information
  - name: Satheesh Joseph
    firstname: Satheesh 
    surname: Joseph
    email: satheeshrishi@berkeley.edu
    affiliation: UC Berkeley, School of Information
acknowledgements: | 
  The authors would like to thank their instructors from MIDS 271.
abstract: | 
  The year is 1997 and global attention is turning toward the consequences of human-actions in our environmental system. The Intergovernmental Panel on Climate Change (IPCC) has been in existence and studying these trends for more than ten years. It has released its second assessment report in 1995. In this report, the IPCC notes that the balance of the evidence suggests that human-actions play a role in the changing climate. It appears that while there are skepticism around the theory of global warming being attributed to human-actions the world hasn't yet been polazrized to be in pro- and anti- global warming camps. That said, there is also little political will to change the current trend contributing to the global warming effects. In this report, we assess data from the Mona Loa Observatory (MLO) to describe and predict global $CO_{2}$ concentrations under several possible scenarios. What we find, when we run the analysis, may paint a grim picture. 
header-includes: 
  - '\usepackage{graphicx}'
  - '\usepackage{booktabs}'
output: rticles::aea_article
---

```{r note}
## Students this file and the supporting files to create 
## the document come from the `rticles` package. 
## 
## You're not required to write something using this template; 
## and, you might think that it hampers understanding. That's 
## totally fine. If you would like to see other tempates, 
## you can install, the `rticles` package, and then use the
## templates it provides. 
## install.packages('rticles')
## would do the trick. 
```

```{r setup, echo=FALSE}
## default to not show code, unless we ask for it.
knitr::opts_chunk$set(echo=FALSE)
options(digits = 3)
```

```{r load packages, echo = FALSE, message = FALSE, include=FALSE}
library(tidyverse)
library(tsibble)
library(latex2exp)
library(ggplot2)
library(patchwork)
library(magrittr)
library(lubridate)
library(feasts)
library(forecast)
library(fable)
library(sandwich)
library(lmtest)
library(gridExtra)
library(scales)
library(zoo)

install.packages("astsa")
library(astsa)

theme_set(theme_minimal())
knitr::opts_chunk$set(dpi=1000)
```

Understanding a changing climate, and what it means for the earth's inhabitants is of growing interest to the scientific and policy community. Although, at this point in 1997 it is not entirely clear what the consequences of this growing awareness will be, in this report we present likely outcomes under "business-as-usual" scenarios. In doing so, our hope, is to establish a series of possible futures, and, as evidence, technology, and policy develop over the coming decades, that we can weigh the impacts that carbon-emission reduction efforts might take. 

```{r plot the keeling curve, echo = FALSE}
tsibble::as_tsibble(co2) %>%
  ggplot() + 
  aes(x=index, y=value) + 
  geom_line(color = 'steelblue') +
  labs(
    title = TeX(r'(Monthly Mean $CO_2$)'),
    subtitle = 'The "Keeling Curve"',
    x = 'Month and Year',
    y = TeX(r'($CO_2$ parts per million)')
  )
```
# Task 0a: Introduction


# Task 1a: EDA

```{r generate the required data object, echo=FALSE}
# generate tsibble object that's most conducive for analysis
co2_tsb <- co2 %>% as_tsibble()
colnames(co2_tsb)[2] <- "co2_ppm"
head(co2_tsb)

# check for missing values
num_na <- sum(is.na(co2_tsb$co2_ppm))
if (num_na > 0) {
  print(paste("There are", num_na, "missing values"))
} else {
    print("There are no missing values in the measurement of CO2 levels") 
}
```

```{r Histogram}
hist(co2, breaks = 30, col="gray", freq = FALSE, 
     xlab = "CO2 ppm", main = "Histogram of the CO2 Concentration")
lines(density(co2), col = 'red', lty = 3)
leg.txt <- c("Estimated density plot")
legend("topright", legend = leg.txt, lty = 2, col = "red", bty = 'n', cex = .8)
```

```{r Time Series Plot}
# Check whether time-series is mean stationary

plot.ts(co2_tsb$co2_ppm, col = 'red', type = 'l', 
     xlab = "Year (time period: month)", ylab = "CO2 ppm", 
     main = "Time-series plot of the CO2 concentration")
abline(h = mean(co2), col = 'green', lty = 2)
lines(stats::filter(co2_tsb$co2_ppm, sides=2, rep(1, 12)/12), lty = 1, lwd = 1.5, 
      col = "blue")
leg.txt <- c("Time-series", "Mean value", 
             "12-Month Symmetric Moving Average")
legend("topleft", legend = leg.txt, lty = c(1, 2, 1), lwd = c(1, 1, 1.5), 
       col = c("red", "green", "blue"), bty = 'n', cex = .8)
```

```{r Boxplot}
# Check whether time-series is variation stationary

boxplot(co2 ~ factor(rep(1959:1997, each = 12)), 
        xlab = 'Year', ylab = 'CO2 PPM',
        outcex = 0.8, medcol="blue", lwd = 0.3, 
        main = 'Annual Variation of CO2 Concentration')
```

```{r Decomposition}
co2_tsb %>%
  model(STL(co2_ppm ~ trend() + season(),
            robust = TRUE)) %>%
  components() %>%
  autoplot()
```


```{r EDA to see the time-series and its basic characteristics, echo=FALSE}
# co2_tsb %>% gg_tsdisplay(co2_ppm, plot_type="partial")
co2_tsb %>% 
  gg_season(co2_ppm) +
  ylab("CO2 ppm") +
  ggtitle("Annual Seasonality")
co2_tsb %>%
  gg_subseries(co2_ppm) +
  ylab("CO2 ppm") +
  ggtitle("Monthly Trend")
acf(co2_tsb, lag.max = 48, main = "ACF plot")
pacf(co2_tsb, lag.max = 48, main = "PACF plot")
```

# Task 2a: Linear time trend model


```{r Unit Root Test}
# Check stationarity of original data using Unitroot test
co2_tsb %>%
  features(co2_ppm, unitroot_kpss)

```

Original time-series clearly reflects seasonality as well as trend.
Unit root test p-value = 0.01 (< 0.05) is significant at 5% significance level, confirming that we can reject the null hypothesis that original data is stationary. ACF plot indicates 'ANNUAL' seasonality which is also confirmed by PACF at 12th lag. 

We apply differencing at 12th lag to remove seasonality and retest updated time-series using unit root test.
```{r Differencing at 12th lag}
# Apply differencing at 12th lag to remove annual seasonality
co2_annual<- co2 %>%
  diff(lag=12) %>%
  ggtsdisplay(lag.max=48)
```
```{r Differencing Histogram}
co2_tsb %>%
  ggplot(aes(x=difference(co2_ppm, 12))) +
  geom_histogram(binwidth = 0.10, color="black", fill="grey") +
  labs(x = 'Annual season differencing')
```

```{r Unit Root Test}
# Check stationarity of 1st order (12 month seasonality) differencing data using Unitroot test
co2_tsb %>%
  features(difference(co2_ppm, 12), unitroot_kpss)
```

Differenced time-series DOES NOT appear stationary as sinusoidal pattern is clearly visible.
Histogram shows that differenced data is fairly normal, indicating that differencing data is fairly centered around mean.
Unit Root test p-value = 0.01 (< 0.05) is STILL significant at 5% significance level, confirming we reject the null hypothesis that 1st order differenced data is stationary. ACF confirms the same with slow tapering down followed by very slow tapering up. PACF also confirms residual seasonality with sinusoidal behavior with correlation values crossing the significance bounds on either sides.

```{r 2nd order differencing of 1 lag}
# Apply 2nd order differencing of 1 lag to check stationarity
co2_annual_plus1<- co2 %>%
  diff(lag=12) %>%
  diff(lag=1) %>%
  ggtsdisplay(lag.max=48)
```
```{r}
co2_tsb %>%
  ggplot(aes(x=difference(difference(co2_ppm, 12),1))) +
  geom_histogram(binwidth = 0.10, color="black", fill="grey") +
  labs(x = '2nd order differencing')
```

```{r}
co2_tsb %>%
  features(difference(difference(co2_ppm, 12), 1), unitroot_kpss)
```
Differenced of difference time-series looks reasonably stationary around mean of 0.
Histogram of 2nd order differencing is normally distributed and gives confidence that data is centered around mean or in other words, stationary.
Unit root test p-value of 0.1 (>0.05) confirms that series is now stationary. ACF plots still shows out of significance bound correlations upto 12th lag but follow up lags are well within bounds. PACF also reflects sinusoidal pattern indicating residual seasonality.
Overall, we can proceed with model building, as modified time series seems reasonably stationary based upon p-value and we would prefer not to over fit our model.


## Linear Model
```{r Linear Model}
co2_1997_linear <- co2_tsb %>%
  model(TSLM(co2_ppm ~ trend()))

report(co2_1997_linear)
```

```{r Linear Model and Original Time Series plot}
augment(co2_1997_linear) %>%
  ggplot(aes(x=index)) +
  geom_line(aes(y=.fitted, colour = 'LinearFit')) +
  geom_line(aes(y=co2_ppm, colour = 'Original')) +
  labs(title='CO2 ppm vs. Linear Model') +
  scale_colour_manual(values = c(LinearFit = 'red', Original = 'blue')) +
  guides(colour = guide_legend(title = 'Legend'))
```

Linear fit does a decent job as it stays with original data bounds for the time frame, but it does not guarantee a good forecast.

```{r Residuals of Linear Fit Model}
co2_1997_linear %>% 
  gg_tsresiduals() +
  labs(title = "Residuals of a Linear Model fit")
```

There is a strong correlation and seasonality in residuals as reflected in ACF as well. Residuals are not centered around a mean either mean goes down and then goes up over time. Histogram reflects the same behavior.


```{r Logarithmic Transformation}
co2_1997_log <- co2_tsb %>%
  model(TSLM(log(co2_ppm) ~ trend()))

report(co2_1997_log)
```

```{r Logarithm Model and Original Time Series plot}
augment(co2_1997_log) %>%
  ggplot(aes(x=index)) +
  geom_line(aes(y=.fitted, colour = 'LogFit')) +
  geom_line(aes(y=co2_ppm, colour = 'Original')) +
  labs(title='CO2 ppm vs. Logarithmic Model') +
  scale_colour_manual(values = c(LogFit = 'red', Original = 'blue')) +
  guides(colour = guide_legend(title = 'Legend'))
```

```{r Residuals of Linear Fit Model}
co2_1997_log %>% 
  gg_tsresiduals() +
  labs(title = "Residuals of a Logarithmic Model fit")

```
### WRITE UP for LOG Transform and Residual Analysis

```{r Quadratic Model}
co2_1997_quad <- co2_tsb %>%
  model(TSLM(co2_ppm ~ trend() + I(trend()^2)))

report(co2_1997_quad)
```

```{r Logarithm Model and Original Time Series plot}
augment(co2_1997_quad) %>%
  ggplot(aes(x=index)) +
  geom_line(aes(y=.fitted, colour = 'QuadFit')) +
  geom_line(aes(y=co2_ppm, colour = 'Original')) +
  labs(title='CO2 ppm vs. Logarithmic Model') +
  scale_colour_manual(values = c(QuadFit = 'red', Original = 'blue')) +
  guides(colour = guide_legend(title = 'Legend'))

```

```{r Residuals of Quad Fit Model}
co2_1997_quad %>% 
  gg_tsresiduals() +
  labs(title = "Residuals of a Quadratic Model fit")
```

```{r Quadratic_2 Model}
co2_1997_quad_2 <- co2_tsb %>%
  model(TSLM(co2_ppm ~ trend() + I(trend()^2) + I(trend()^3)))

report(co2_1997_quad_2)
```

```{r Logarithm Model and Original Time Series plot}
augment(co2_1997_quad_2) %>%
  ggplot(aes(x=index)) +
  geom_line(aes(y=.fitted, colour = 'Quad2Fit')) +
  geom_line(aes(y=co2_ppm, colour = 'Original')) +
  labs(title='CO2 ppm vs. Logarithmic Model') +
  scale_colour_manual(values = c(Quad2Fit = 'red', Original = 'blue')) +
  guides(colour = guide_legend(title = 'Legend'))

```

```{r Residuals of Quad Fit Model}
co2_1997_quad_2 %>% 
  gg_tsresiduals() +
  labs(title = "Residuals of a Quadratic Model fit")
```
